% Compile with: pdflatex research_paper.tex
% PDF build was not executed in this shell because pdflatex is not installed here.
% Replace the author names, affiliation, and email addresses before submission.
% Suggested title is original and intentionally different from the cited base paper.

\documentclass[conference]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{url}
\usepackage{multirow}

\begin{document}

\title{Research-Grade Adaptive Hierarchical Federated Learning for 6G Edge Intelligence with Baselines, Fairness, and Security-Aware Evaluation}

\author{
\IEEEauthorblockN{Chetan Ahuja, Darshit Khandelwal, Jay Hirapara, Kundan Meena, Shubham Sharma}
\IEEEauthorblockA{
Department of Information Technology \\
ABV-IIITM , Gwalior, India 
}
}

\maketitle

\begin{abstract}
Sixth-generation (6G) wireless systems are expected to support both massive device populations and services with stringent reliability and timeliness requirements. These demands create pressure on the learning, communication, and orchestration layers of future edge-native networks. Motivated by recent 6G connectivity research, this paper presents an implementation-oriented study of an adaptive hierarchical federated learning (FL) framework designed for a three-tier device-edge-cloud architecture. The proposed framework combines fairness-aware edge-side client selection, quality-aware weighted aggregation, hybrid anomaly screening based on both update norms and update directions, and global model synchronization across rounds. Unlike flat FL, the cloud does not aggregate every selected client update directly; instead, regional edge servers perform intermediate aggregation and transmit only validated regional models. The artifact now supports a benchmark suite with flat and hierarchical baselines, five-seed evaluation, real datasets such as Digits and optional torchvision benchmarks, and exports for communication, latency, energy, convergence, fairness, class-balance, and attack-screening metrics. On the clean Digits benchmark, the strongest benign-only result is achieved by the fairness-aware no-filter ablation at 0.8872 final accuracy and 0.8847 macro-F1, while the proposed method attains 0.8644 accuracy with similarly strong participation fairness (0.8985 Jain fairness and 0.1797 Gini). Under a 25\% sign-flip attack benchmark, the proposed method becomes the best overall operating point, reaching 0.8756 final accuracy, 0.8697 macro-F1, 100\% attack detection, 100\% benign retention, and the strongest fairness score among all methods. The resulting contribution is therefore not a claim of universal dominance on every benign metric, but a benchmark-oriented software artifact that exposes the trade-off between clean-set efficiency and security-aware reliability in hierarchical FL for 6G-inspired edge environments.
\end{abstract}

\begin{IEEEkeywords}
6G, federated learning, hierarchical federated learning, edge intelligence, client selection, fairness, robust aggregation, anomaly filtering, benchmarking, wireless networks
\end{IEEEkeywords}

\section{Introduction}
\IEEEPARstart{T}{he} transition from 5G to 6G is not only a matter of higher throughput, but also of architectural adaptation to dense device populations, distributed intelligence, and service classes with strict reliability and timeliness constraints. Recent work by Kal{\o}r \emph{et al.} argues that future 6G systems must jointly address massive connectivity, mixed-criticality traffic, and distributed intelligence, with decentralized learning and inference acting as important drivers of closed-loop traffic patterns and edge-native coordination \cite{kalor2025wireless6g}. This is especially relevant in Internet-of-Things (IoT), cyber-physical, industrial, and autonomous environments, where many devices must collaborate without violating privacy or overwhelming backhaul resources.

Federated learning (FL) is attractive because it keeps raw data on local devices and exchanges model information instead \cite{mcmahan2017fedavg}. However, classical flat FL has clear systems limitations in wireless environments: not all clients are equally reliable, communication conditions vary significantly, full participation is expensive, and the cloud becomes a bottleneck when every device communicates globally in each round \cite{kairouz2021advances,amiri2020wireless,chen2021joint}. These limitations become more severe in 6G scenarios that mix high device density, edge intelligence, and service-critical operation.

This paper presents a research-oriented upgrade of our hierarchical FL artifact. Rather than treating FL as a flat device-to-cloud procedure, we build a benchmark-capable device-edge-cloud workflow in which edge servers perform intermediate intelligence before cloud synchronization. The contribution is therefore not only a working hierarchical FL implementation, but also an evaluation-ready software framework that supports direct baselines, ablation studies, fairness analysis, and security-aware reporting.

It is also important to state the privacy scope precisely. The present system improves privacy relative to centralized learning by keeping raw data on devices and exchanging only model information. However, it does not yet implement formal privacy-enhancement mechanisms such as differential privacy, secure aggregation, or cryptographic protection of model updates.

The main contributions of this paper are as follows:
\begin{itemize}
    \item We design a research-grade three-tier hierarchical FL framework in which device clients train locally, edge servers perform fairness-aware intelligent client selection and regional aggregation, and the cloud performs final global synchronization.
    \item We implement direct comparison baselines and ablation settings, including flat FedAvg, hierarchical random selection, intelligent selection without weighted aggregation, weighted aggregation without filtering, and weighted aggregation without fairness balancing.
    \item We extend the evaluation pipeline with communication, latency, energy, convergence, fairness, class-balance, and attack-screening metrics, together with exportable benchmark reports and \LaTeX{} summary tables.
    \item We support both synthetic non-IID data and benchmark datasets such as Digits, with optional torchvision-backed MNIST, Fashion-MNIST, and CIFAR-10 evaluation \cite{darshit2026repo}.
\end{itemize}

The remainder of the paper is organized as follows. Section II discusses the motivation and related work. Section III presents the system model. Section IV details the proposed adaptive hierarchical FL framework. Section V describes the implementation and benchmark setup. Section VI discusses result interpretation and publication reporting. Section VII outlines limitations and future work, and Section VIII concludes the paper.

\section{Background, Motivation, and Related Work}
\subsection{6G Connectivity Perspective}
The base paper in \cite{kalor2025wireless6g} argues that 6G must be understood not only as a faster wireless interface, but as a system-level response to two intertwined demands: connectivity for massive numbers of devices and connectivity for critical services. Massive connectivity involves scaling beyond conventional mobile-user assumptions, while critical services require extremely reliable and timely operation under diverse timing semantics. The same work further notes that distributed and decentralized learning can reshape traffic patterns by increasing the amount of closed-loop and symmetric communication in future networks \cite{kalor2025wireless6g}. This creates a strong motivation for edge-resident intelligence rather than purely centralized coordination.

\subsection{Federated Learning Over Wireless Networks}
FedAvg introduced the modern FL paradigm by showing that decentralized data holders can collaboratively train a global model using iterative local updates and centralized averaging \cite{mcmahan2017fedavg}. Since then, a large body of work has examined communication, privacy, optimization, and systems challenges in FL \cite{kairouz2021advances}. Within wireless networking, research has addressed fading-channel effects, over-the-air aggregation, resource allocation, convergence under limited bandwidth, and energy-aware participation \cite{amiri2020wireless,chen2021joint,yang2021energy}. Hierarchical FL has also been studied explicitly in the client-edge-cloud setting. In particular, Liu \emph{et al.} proposed a HierFAVG-style architecture in which edge servers perform intermediate aggregation to improve training speed and communication-computation trade-offs \cite{liu2020hierfavg}.

Despite this progress, many implementations still assume a flat topology in which a base station or cloud server directly coordinates all users. Such designs are difficult to scale to very large client populations and are also vulnerable to poor client selection. In practice, many clients may have insufficient battery, poor latency, or limited data value, yet they may still participate if the selection policy is naive. For 6G-inspired deployments, this motivates a hierarchical architecture in which edge nodes filter, select, and aggregate before the cloud performs cross-region synchronization.

\subsection{Gap Addressed by This Work}
This paper addresses a specific implementation gap. The literature provides strong theoretical motivation for 6G-compatible distributed learning, but there is still a need for transparent and reproducible software artifacts that support actual comparative experiments. Our work therefore focuses on the orchestration layer of hierarchical FL and turns a functional prototype into a benchmark-oriented research artifact rather than a one-off demonstration.

\section{System Model and Problem Formulation}
\subsection{Three-Tier Network Architecture}
We consider a hierarchical FL system with three layers:
\begin{itemize}
    \item \textbf{Device layer:} A set of client devices $\mathcal{C}$ stores local data and performs local learning.
    \item \textbf{Edge layer:} A set of edge servers $\mathcal{E}$ coordinates nearby clients, selects participants, filters suspicious updates, and performs regional aggregation.
    \item \textbf{Cloud layer:} A global server receives regional edge models and produces the next global model.
\end{itemize}

Let $\mathcal{C}_e \subset \mathcal{C}$ denote the clients assigned to edge server $e$. At communication round $t$, the cloud holds model weights $\mathbf{w}^{(t-1)}$, which are transmitted to the edge layer and then to the selected clients. Each participating client trains locally and returns an update to its associated edge server. The edge server aggregates local updates into a regional model $\mathbf{w}_e^{(t)}$, and the cloud combines all regional models to form the next global model $\mathbf{w}^{(t)}$.

\subsection{Device State Representation}
Each client $i$ is characterized by a state vector
\begin{equation}
\mathbf{s}_i^{(t)} = \left[b_i^{(t)}, \ell_i^{(t)}, n_i, d_i, r_i, f_i^{(t)} \right],
\end{equation}
where $b_i^{(t)}$ denotes battery level, $\ell_i^{(t)}$ denotes network latency, $n_i$ denotes local data size, $d_i$ denotes a data-diversity score, $r_i$ denotes a reliability score, and $f_i^{(t)}$ denotes a freshness score that reflects how long the client has waited since its last selection. In the implementation, battery and latency evolve across rounds to emulate realistic variation in device conditions. Specifically,
\begin{align}
b_i^{(t+1)} &= \mathrm{clip}\left(b_i^{(t)} - \delta_i^{(t)}, 15, 100 \right), \\
\ell_i^{(t+1)} &= \mathrm{clip}\left(\ell_i^{(t)} + \xi_i^{(t)}, 5, 200 \right),
\end{align}
where $\delta_i^{(t)}$ is an integer drain in the range $[1,4]$ and $\xi_i^{(t)}$ is a latency perturbation in the range $[-8,8]$.

\subsection{Local Learning Objective}
Each client $i$ minimizes a local empirical risk over its dataset $\mathcal{D}_i$:
\begin{equation}
F_i(\mathbf{w}) = \frac{1}{|\mathcal{D}_i|}\sum_{(\mathbf{x},y)\in\mathcal{D}_i} \mathcal{L}(f(\mathbf{x};\mathbf{w}), y),
\end{equation}
where $f(\cdot;\mathbf{w})$ is the model and $\mathcal{L}(\cdot,\cdot)$ is the cross-entropy loss. The client starts each round from the incoming global model and produces an updated model $\mathbf{w}_i^{(t)}$. Its local delta is
\begin{equation}
\Delta_i^{(t)} = \mathbf{w}_i^{(t)} - \mathbf{w}^{(t-1)}.
\end{equation}

\subsection{Design Objective}
The objective of the system is not merely to optimize local training accuracy, but to do so under constraints relevant to 6G-style massive connectivity:
\begin{itemize}
    \item limit the number of active clients per round,
    \item prefer clients with more favorable operating conditions,
    \item improve fairness of participation over time,
    \item reduce direct cloud aggregation load through edge summarization,
    \item avoid destabilizing the global model with anomalous client updates,
    \item support reproducible direct comparison against baseline methods.
\end{itemize}

\section{Proposed Adaptive Hierarchical FL Framework}
\subsection{Overview}
The proposed framework consists of five tightly coupled stages performed in each round:
\begin{enumerate}
    \item fairness-aware edge-side intelligent client selection,
    \item local client training from the latest global model,
    \item conservative anomaly screening of returned updates,
    \item robust quality-aware edge aggregation,
    \item sample-weighted cloud aggregation.
\end{enumerate}

\subsection{Edge-Side Intelligent Client Selection}
Rather than selecting clients uniformly at random, each edge server ranks the clients in its region using a logistic regression model operating on standardized device-state features. This realizes the idea of edge-native intelligent client selection in a lightweight form suitable for deployment, without requiring the complexity of reinforcement learning or online policy optimization in the current artifact. The base probability of selecting client $i$ at round $t$ is modeled as
\begin{equation}
p_i^{(t)} = \sigma \left(\boldsymbol{\theta}^{\top}\tilde{\mathbf{s}}_i^{(t)} \right),
\end{equation}
where $\tilde{\mathbf{s}}_i^{(t)}$ is the standardized feature vector and $\sigma(\cdot)$ is the sigmoid function.

To initialize the selector without requiring a pre-existing labeled dataset, the implementation first generates synthetic training samples using a rule-based device desirability score
\begin{equation}
\psi_i = 0.28 \frac{b_i}{100} + 0.22\left(1 - \frac{\ell_i}{200}\right) + 0.18 \min\left(\frac{n_i}{1000},1\right) + 0.14 r_i + 0.10 d_i + 0.08 f_i.
\end{equation}
Binary labels are then assigned using a threshold $\tau = 0.58$:
\begin{equation}
y_i =
\begin{cases}
1, & \psi_i \ge \tau, \\
0, & \text{otherwise}.
\end{cases}
\end{equation}
This creates a lightweight surrogate for edge-side admission control. In each round, the final ranking score is adjusted with diversity, reliability, freshness, participation gap, and first-time coverage bonuses so that repeatedly selected strong clients do not dominate participation indefinitely. In the fairness-aware mode, selection is not a strict deterministic top-$K$ rule; instead, clients are sampled without replacement with probabilities proportional to their adjusted scores. In this sense, the upgraded selector is not only utility-aware, but also participation-aware.

\subsection{Synthetic and Benchmark Non-IID Client Data}
The artifact supports both synthetic and benchmark datasets. For the synthetic setting, client-specific datasets emulate statistical heterogeneity. For each client, feature vectors are sampled around a client-specific random center, and labels are generated using
\begin{equation}
y = \mathbb{I}\left(0.9x_1 - 0.6x_2 + 0.4x_3 + 0.3x_4 + \epsilon > 0 \right),
\end{equation}
where $\epsilon$ is Gaussian noise and $\mathbb{I}(\cdot)$ is the indicator function. This creates non-identical data distributions across clients without requiring an external dataset. In addition, the artifact supports benchmark datasets such as Digits and optional torchvision datasets using Dirichlet non-IID partitioning across clients. This makes the system suitable for both self-contained demonstrations and stronger publication-style evaluation.

\subsection{Local Training}
Each selected client receives the current global model and performs a small number of local training epochs. The model used in this study is a compact multilayer perceptron whose input dimension adapts to the active dataset. The architecture follows the pattern:
\begin{itemize}
    \item dataset-dependent input dimension,
    \item one hidden layer with configurable width (default $32$),
    \item ReLU activation,
    \item output dimension matched to the dataset class count.
\end{itemize}
The client returns its trained model, the update delta, local accuracy, local loss, and metadata used by the edge aggregator, including upload size, train-time proxy, energy proxy, and attack status.

\subsection{Quality-Aware Edge Aggregation}
To avoid treating all clients equally, each client receives a transmitted quality score that combines system condition, data diversity, historical utility, and immediate local-learning quality. A normalized system-condition score is defined as
\begin{equation}
u_i^{(t)} = 0.35 \frac{b_i^{(t)}}{100} + 0.20\left(1 - \frac{\ell_i^{(t)}}{200}\right) + 0.30 \min\left(\frac{n_i}{900},1\right) + 0.15 r_i.
\end{equation}
The immediate learning-quality score is
\begin{equation}
\lambda_i^{(t)} = 0.45 a_i^{(t)} + 0.25 e^{-L_i^{(t)}} + 0.20 d_i + 0.10 h_i^{(t)},
\end{equation}
where $a_i^{(t)}$ and $L_i^{(t)}$ denote local evaluation accuracy and loss, respectively, and $h_i^{(t)}$ denotes historical utility. The transmitted quality score is then
\begin{equation}
q_i^{(t)} = 0.30 u_i^{(t)} + 0.45 \lambda_i^{(t)} + 0.15 d_i + 0.10 r_i.
\end{equation}
This score is combined with the local data size to define the aggregation weight
\begin{equation}
\alpha_i^{(t)} = \frac{n_i q_i^{(t)}}{\sum_{j \in \mathcal{S}_e^{(t)}} n_j q_j^{(t)}},
\end{equation}
where $\mathcal{S}_e^{(t)}$ is the set of safe selected clients associated with edge server $e$ in round $t$. The edge model is then computed as
\begin{equation}
\mathbf{w}_e^{(t)} = \mathbf{w}^{(t-1)} + \sum_{i \in \mathcal{S}_e^{(t)}} \alpha_i^{(t)} \Delta_i^{(t)}.
\end{equation}

\subsection{Conservative Anomaly Screening}
Security is handled conservatively. For each client update, the edge server computes both a norm over the flattened delta vector and a directional similarity score against the edge consensus update. A robust median absolute deviation (MAD) filter is then used to identify abnormal update norms:
\begin{equation}
z_i = \frac{\left|\|\Delta_i^{(t)}\| - \mathrm{median}(\mathcal{N})\right|}{1.4826 \cdot \mathrm{MAD}(\mathcal{N})},
\end{equation}
where $\mathcal{N}$ denotes the set of delta norms at the edge. In addition, each client update is compared with the coordinate-wise median direction of the edge batch using cosine similarity. Updates are removed if they are either norm outliers or directionally inconsistent with the edge consensus. This hybrid rule is especially useful for attacks such as sign flipping, where malicious updates may point in the opposite direction even when their norms alone are not sufficient to trigger rejection. However, the filter is applied only when at least four client updates are available at the edge, which prevents unreliable decisions from very small samples. The upgraded benchmark suite further tracks blocked adversarial updates, accepted adversarial updates, filter precision, and benign retention for security-aware reporting.

\subsection{Cloud Aggregation}
Once each edge server produces a regional model, the cloud performs sample-weighted averaging:
\begin{equation}
\mathbf{w}^{(t)} = \sum_{e \in \mathcal{E}} \beta_e^{(t)} \mathbf{w}_e^{(t)},
\end{equation}
with
\begin{equation}
\beta_e^{(t)} = \frac{N_e^{(t)}}{\sum_{k \in \mathcal{E}} N_k^{(t)}},
\end{equation}
where $N_e^{(t)}$ is the total number of safe client samples contributing to edge $e$ in round $t$. If an entire edge region is filtered out in a round, the cloud safely skips that region instead of failing.

\section{Implementation and Experimental Setup}
\subsection{Software Stack}
The framework is implemented in Python using PyTorch for local learning, NumPy for numerical operations, scikit-learn for classical benchmark data utilities and the lightweight logistic-regression-based selector, pandas for result export, and matplotlib for visualization. The complete source code is publicly available in the project repository \cite{darshit2026repo}.

\subsection{Benchmark Configuration}
Table \ref{tab:setup} summarizes the exact benchmark configuration used for the main Digits results reported in this paper. The same core setup is reused for both the clean benchmark and the sign-flip attack benchmark, with the attack parameters enabled only for the latter.

\begin{table}[t]
\caption{Benchmark Setup Used in This Paper}
\label{tab:setup}
\centering
\small
\begin{tabular}{>{\raggedright\arraybackslash}p{3.2cm}p{4.7cm}}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Seeds & 42, 52, 62, 72, 82 \\
Total number of clients & 20 \\
Number of edge servers & 2 \\
Selected clients per edge per round & 4 \\
Total communication rounds & 10 \\
Local epochs & 2 \\
Learning rate & 0.02 (default) \\
Hidden width & 32 \\
Reported dataset & Digits \\
Partitioning & Dirichlet non-IID split with $\alpha = 0.5$ \\
Selection modes & Random, intelligent, fairness-aware intelligent \\
Aggregation modes & FedAvg, quality-weighted aggregation \\
Attack models supported & Sign-flip, Gaussian-noise, label-flip, model-replacement \\
Reported attack benchmark & 25\% sign-flip attack, scale $5.0$ \\
Security filter & Hybrid norm/direction filter (MAD $3.5$, cosine MAD $3.0$) \\
Minimum updates for filtering & 4 \\
Exports & CSV, JSON, markdown report, \LaTeX{} table, plots \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Evaluation Metrics}
The main evaluation metrics are:
\begin{itemize}
    \item global model accuracy after each communication round,
    \item global model loss after each communication round,
    \item macro-F1, balanced accuracy, and worst-class accuracy,
    \item communication payloads to the edge and cloud layers,
    \item total communication bytes,
    \item latency and energy proxy metrics,
    \item train-time proxy metrics,
    \item convergence speed and fairness of client participation,
    \item participation entropy, coverage, and Gini inequality,
    \item attack-screening metrics such as blocked malicious updates, attack escape rate, filter precision, and benign retention.
\end{itemize}

The benchmark artifact now includes a flat FedAvg baseline, hierarchical ablations that remove one system component at a time, repeated multi-seed execution, explicit adversarial update injection through sign-flip, Gaussian-noise, label-flip, and model-replacement attacks, and per-client participation exports. The focus therefore shifts from ``showing that the system works'' to enabling direct comparative evaluation.

\section{Results and Discussion}
\subsection{Clean Benchmark Results}
The upgraded artifact was executed on the Digits benchmark across five seeds using the benchmark configuration in Table \ref{tab:setup}. Table \ref{tab:results} reports the main clean-benchmark comparison obtained from the exported summary files in the repository.

\begin{table}[t]
\caption{Main Results on the Clean Digits Benchmark (5 Seeds)}
\label{tab:results}
\centering
\small
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lccccccc}
\toprule
\textbf{Method} & \textbf{Final Acc.} & \textbf{Macro-F1} & \textbf{Bal. Acc.} & \textbf{Worst Cls.} & \textbf{Cloud Up.} & \textbf{Fairness} & \textbf{Gini} \\
\midrule
\texttt{flat\_fedavg\_random} & 0.8717 & 0.8661 & 0.8740 & 0.5746 & 80 & 0.8804 & 0.2027 \\
\texttt{hierarchical\_random\_fedavg} & 0.8533 & 0.8381 & 0.8577 & 0.4288 & 20 & 0.8544 & 0.2310 \\
\texttt{hierarchical\_intelligent\_fedavg} & 0.8750 & 0.8720 & 0.8735 & 0.5833 & 20 & 0.5219 & 0.5227 \\
\texttt{hierarchical\_intelligent\_no\_filter} & \textbf{0.8872} & \textbf{0.8847} & \textbf{0.8846} & \textbf{0.5934} & 20 & \textbf{0.8993} & 0.1802 \\
\texttt{hierarchical\_intelligent\_no\_fairness} & 0.8733 & 0.8645 & 0.8696 & 0.4529 & 20 & 0.5354 & 0.5068 \\
\texttt{proposed} & 0.8644 & 0.8583 & 0.8660 & 0.4319 & 20 & 0.8985 & \textbf{0.1797} \\
\bottomrule
\end{tabular}
}
\end{table}

\subsection{Clean-Benchmark Interpretation}
The clean benchmark reveals a more nuanced trade-off than a simple ``proposed method wins everywhere'' narrative. The strongest benign-only result is obtained by \texttt{hierarchical\_intelligent\_no\_filter}, which reaches 0.8872 final accuracy and 0.8847 macro-F1 while also achieving strong fairness (0.8993). The proposed method remains highly competitive on fairness (0.8985 Jain fairness and the lowest Gini score at 0.1797), but its clean-set accuracy drops to 0.8644. This gap is consistent with the fact that the proposed method applies conservative filtering even in benign conditions. In the clean five-seed run, the proposed method retains 92.25\% of benign updates on average, indicating that a small number of legitimate updates are rejected together with suspicious ones. This is a meaningful cost, and it should be stated explicitly.

At the same time, the clean benchmark still supports two publication-relevant claims. First, fairness-aware hierarchical coordination can maintain very balanced participation while preserving strong predictive quality. Second, the hierarchy reduces cloud-facing uploads from 80 to 20 over the full run, corresponding to a 75\% reduction in cloud fan-in. However, the total payload bytes rise from 771{,}200 in flat FedAvg to 964{,}000 in the hierarchical runs because this simulation counts both device-to-edge and edge-to-cloud transmissions and does not implement compression. Therefore, the present artifact demonstrates backhaul relief and aggregation decentralization rather than absolute end-to-end byte minimization.

\subsection{Metrics That Strengthen Publication Claims}
Compared with the earlier prototype-oriented version of the project, the upgraded artifact supports much stronger publication claims because it measures:
\begin{itemize}
    \item end-to-end predictive performance through accuracy and loss,
    \item class-balance behavior through macro-F1, balanced accuracy, and worst-class accuracy,
    \item communication efficiency through payload scalars and bytes,
    \item latency and compute behavior through round-latency and train-time proxies,
    \item fairness through Jain's fairness index, participation entropy, client coverage, and Gini inequality,
    \item security behavior through attack detection, attack escape rate, filter precision, and benign retention,
    \item convergence behavior through accuracy gain and convergence round.
\end{itemize}

\subsection{Scalability Interpretation}
The structure of the system establishes a clear scalability argument, but the metric interpretation must be precise. Under the same participation budget, the hierarchical design reduces cloud-facing aggregation fan-in because the cloud receives regional edge updates rather than direct uploads from every selected client. In the reported benchmark this reduces cloud uploads by 75\%. However, the current implementation does not compress edge payloads or overlap transmissions, so total payload bytes are higher than flat FedAvg. For this reason, the strongest systems claim supported by the present results is reduced cloud coordination burden, not universal communication optimality.

\subsection{Security Benchmark Under Adversarial Updates}
To evaluate the anomaly-filtering mechanism under attack, we executed a second Digits benchmark with the same five seeds and a 25\% sign-flip attack fraction. The resulting comparison is shown in Table \ref{tab:attack}.

\begin{table}[t]
\caption{Results on the Adversarial Digits Benchmark (25\% Sign-Flip, 5 Seeds)}
\label{tab:attack}
\centering
\small
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lcccccc}
\toprule
\textbf{Method} & \textbf{Final Acc.} & \textbf{Macro-F1} & \textbf{Bal. Acc.} & \textbf{Worst Cls.} & \textbf{Fairness} & \textbf{Atk. Detect} \\
\midrule
\texttt{flat\_fedavg\_random} & 0.1011 & 0.0183 & 0.1000 & 0.0000 & 0.8804 & 0.0000 \\
\texttt{hierarchical\_random\_fedavg} & 0.0917 & 0.0168 & 0.1000 & 0.0000 & 0.8485 & 0.0000 \\
\texttt{hierarchical\_intelligent\_fedavg} & 0.1017 & 0.0186 & 0.0994 & 0.0000 & 0.5763 & 0.0000 \\
\texttt{hierarchical\_intelligent\_no\_filter} & 0.1150 & 0.0279 & 0.1161 & 0.0000 & 0.8957 & 0.0000 \\
\texttt{hierarchical\_intelligent\_no\_fairness} & 0.8422 & 0.8295 & 0.8414 & 0.3868 & 0.5980 & 1.0000 \\
\texttt{proposed} & \textbf{0.8756} & \textbf{0.8697} & \textbf{0.8748} & \textbf{0.4962} & \textbf{0.9047} & \textbf{1.0000} \\
\bottomrule
\end{tabular}
}
\end{table}

The adversarial benchmark is especially revealing. Every method without filtering collapses to near-chance behavior, with macro-F1 scores below 0.03. In contrast, the two filtered weighted methods remain highly effective. Among them, the proposed method is the strongest overall operating point, reaching 0.8756 final accuracy, 0.8697 macro-F1, 0.4962 worst-class accuracy, 100\% attack detection, 0\% attack escape, 100\% filter precision, and 100\% benign retention. The variant without fairness also detects all attacks, but it is weaker than the proposed method on every reported predictive metric and substantially weaker on participation fairness (0.5980 versus 0.9047). This result shows that conservative filtering is not merely decorative: it materially changes the behavior of the system under malicious updates, and the fairness-aware selection layer remains useful even when robustness is already in place.

\subsection{Limits of the Reported Metrics}
At the same time, latency, communication load, and energy are still measured through software proxies rather than a detailed wireless simulator or real hardware instrumentation. The reported learning results are also currently based on the Digits benchmark rather than larger-scale vision tasks. Therefore, these metrics should be interpreted as comparative systems indicators rather than exact deployment measurements, and the present conclusions should be read as benchmark evidence for orchestration-layer behavior rather than a complete validation of 6G networking performance. This is still a meaningful improvement over the previous prototype, because the artifact now supports direct method-to-method comparison instead of only descriptive discussion.

\subsection{Relevance to Massive Devices and Critical Services}
The codebase does not directly model queue tails, radio-block scheduling, packet errors, information freshness, or hard latency bounds for critical-control traffic. Therefore, it should not be presented as a complete end-to-end realization of critical-service 6G networking. However, it does implement an important enabling layer for that broader vision: learning orchestration under resource variability and large client populations.

From that perspective, the framework contributes in three ways:
\begin{itemize}
    \item it reduces direct coordination burden at the cloud,
    \item it favors clients with operationally better conditions,
    \item it adds a first layer of update validation before regional aggregation.
\end{itemize}
These properties are directly relevant to 6G-style environments in which distributed learning may coexist with mixed-criticality service classes \cite{kalor2025wireless6g}.

\section{Limitations and Future Research Directions}
Although the artifact is now benchmark-oriented rather than merely prototype-oriented, several limitations remain.

First, the wireless layer is modeled through latency and energy proxies rather than a detailed radio, queueing, or scheduling simulator. Second, the client selector is still a lightweight logistic-regression-based scorer rather than a reinforcement-learning or contextual-bandit policy. Third, the security evaluation currently supports multiple poisoning hooks and a hybrid robust filter, but not a full Byzantine benchmark suite with adaptive attackers or certified defenses. Fourth, the system still does not implement secure aggregation, differential privacy, or cryptographic protection of updates. Fifth, asynchronous participation and client dropouts are not yet modeled in full detail. Sixth, the hierarchy presently reduces cloud fan-in rather than total end-to-end payload bytes, because the implementation does not yet apply compression or communication co-design across the edge hop.

These limitations suggest several future extensions:
\begin{itemize}
    \item stronger security benchmarks with additional attack families,
    \item secure aggregation and formal privacy extensions,
    \item asynchronous or event-driven participation,
    \item richer wireless-resource modeling,
    \item larger topology sweeps and more extensive real-dataset evaluation,
    \item fairness-aware optimization objectives beyond heuristic balancing.
\end{itemize}

\section{Conclusion}
This paper presented a benchmark-oriented upgrade of an adaptive hierarchical federated learning system for 6G-inspired edge intelligence. The upgraded artifact goes beyond a functional prototype by integrating benchmark baselines, fairness-aware client selection, quality-weighted aggregation, anomaly screening, real-dataset support, adversarial testing hooks, and publication-oriented metric exports.

The empirical results show a clear trade-off. In benign conditions, the no-filter fairness-aware ablation achieves the highest clean accuracy, indicating that conservative filtering can reduce peak clean performance when attacks are absent. Under adversarial conditions, however, the proposed method becomes the strongest overall operating point: it combines high predictive quality, perfect attack detection in the reported sign-flip benchmark, full benign retention, and the strongest participation fairness among all methods. The central contribution is therefore not only a hierarchical FL mechanism, but a complete comparative evaluation framework that enables stronger and more credible research claims about the interaction among edge hierarchy, fairness, and robustness in 6G-inspired distributed learning.

\begin{thebibliography}{99}

\bibitem{kalor2025wireless6g}
A.~E. Kal{\o}r, G.~Durisi, S.~Coleri, S.~Parkvall, W.~Yu, A.~Mueller, and P.~Popovski,
``Wireless 6G Connectivity for Massive Number of Devices and Critical Services,''
\emph{Proceedings of the IEEE}, vol.~113, no.~9, pp.~826--848, 2025, doi: 10.1109/JPROC.2024.3484529.

\bibitem{mcmahan2017fedavg}
H.~B. McMahan, E.~Moore, D.~Ramage, S.~Hampson, and B.~A.~y. Arcas,
``Communication-Efficient Learning of Deep Networks from Decentralized Data,''
in \emph{Proc. 20th Int. Conf. Artificial Intelligence and Statistics (AISTATS)}, Fort Lauderdale, FL, USA, 2017, pp.~1273--1282.

\bibitem{kairouz2021advances}
P.~Kairouz \emph{et al.},
``Advances and Open Problems in Federated Learning,''
\emph{Foundations and Trends in Machine Learning}, vol.~14, no.~1--2, pp.~1--210, 2021.

\bibitem{amiri2020wireless}
M.~M. Amiri and D.~G{\"u}nd{\"u}z,
``Federated Learning Over Wireless Fading Channels,''
\emph{IEEE Transactions on Wireless Communications}, vol.~19, no.~5, pp.~3546--3557, May~2020, doi: 10.1109/TWC.2020.2974748.

\bibitem{chen2021joint}
M.~Chen, Z.~Yang, W.~Saad, C.~Yin, H.~V. Poor, and S.~Cui,
``A Joint Learning and Communications Framework for Federated Learning Over Wireless Networks,''
\emph{IEEE Transactions on Wireless Communications}, vol.~20, no.~1, pp.~269--283, Jan.~2021, doi: 10.1109/TWC.2020.3024629.

\bibitem{yang2021energy}
Z.~Yang, M.~Chen, W.~Saad, C.~S. Hong, and M.~Shikh-Bahaei,
``Energy Efficient Federated Learning Over Wireless Communication Networks,''
\emph{IEEE Transactions on Wireless Communications}, vol.~20, no.~3, pp.~1935--1949, Mar.~2021, doi: 10.1109/TWC.2020.3037554.

\bibitem{liu2020hierfavg}
L.~Liu, J.~Zhang, S.~H. Song, and K.~B. Letaief,
``Client-Edge-Cloud Hierarchical Federated Learning,''
in \emph{Proc. IEEE International Conference on Communications (ICC)}, Dublin, Ireland, 2020, Art. no.~9148862, doi: 10.1109/ICC40277.2020.9148862.

\bibitem{darshit2026repo}
Darshit2308,
``wct-6g-federated-learning,''
GitHub repository, 2026. [Online]. Available: \url{https://github.com/darshit2308/wct-6g-federated-learning}

\end{thebibliography}

\end{document}
